---
title: "Optimization of Support Staffing at Tesla"
author: "Valli Kasivishwanath"
date: "2023-12-08"
output: html_document
---

Questions

a)How many full-time English-speaking agents, full-time Spanish-speaking agents, and
part-time agents should Alex hire for each 2-hour shift to minimize operating costs
while attending to all calls? 
Full-Time English-Speaking Agents Needed for Each 2-Hour Shift:

7 A.M. – 9 A.M.: 2 agents, 
 9 A.M. – 11 A.M.: 4 agents, 
 11 A.M. – 1 P.M.: 3 agents, 
 1 P.M. – 3 P.M.: 4 agents, 
 3 P.M. – 5 P.M.: 4 agents,
 5 P.M. – 7 P.M.: 2 agents,
 7 P.M. – 9 P.M.: 1 agent.

Full-Time Spanish-Speaking Agents Needed for Each 2-Hour Shift:

7 A.M. – 9 A.M.: 2 agents
 9 A.M. – 11 A.M.: 4 agents
 11 A.M. – 1 P.M.: 3 agents
 1 P.M. – 3 P.M.: 4 agents
 3 P.M. – 5 P.M.: 3 agents
 5 P.M. – 7 P.M.: 1 agent
 7 P.M. – 9 P.M.: 0 agents

b)What is the minimum cost for the optimization model to assist Alex’s decision in
hiring all agents that she needs? 

```{r Part a,b, echo = FALSE, message=FALSE, comment = " " ,warning=FALSE}
# Define the hourly call forecast data
call_forecast <- c(40, 85, 70, 95, 80, 35, 10)

# Agent efficiency (calls per hour)
agent_efficiency <- 6

# Calculate the number of agents needed for each 2-hour shift
# Full-time agents work effectively for half their shift, part-time agents for their entire shift
full_time_agents_needed <- ceiling(call_forecast / agent_efficiency / 2)
part_time_agents_needed <- c(ceiling(sum(call_forecast[5:6]) / agent_efficiency),
                             ceiling(sum(call_forecast[6:7]) / agent_efficiency))

# Splitting full-time agents into English and Spanish speakers
full_time_english_agents <- ceiling(full_time_agents_needed / 2)
full_time_spanish_agents <- floor(full_time_agents_needed / 2)

# Wage rates
wage_before_5pm <- 30
wage_after_5pm <- 45

# Calculate the cost for full-time and part-time agents
cost_full_time <- sum(full_time_english_agents[1:4] * wage_before_5pm * 2) +
  sum(full_time_spanish_agents[1:4] * wage_before_5pm * 2) +
  sum(full_time_english_agents[5:7] * wage_after_5pm * 2) +
  sum(full_time_spanish_agents[5:7] * wage_after_5pm * 2)

cost_part_time <- part_time_agents_needed[1] * wage_before_5pm * 4 +
  part_time_agents_needed[2] * wage_after_5pm * 4

# Total cost
total_cost <- cost_full_time + cost_part_time

# Output the results
list(
  full_time_english_agents = full_time_english_agents,
  full_time_spanish_agents = full_time_spanish_agents,
  part_time_agents_needed = part_time_agents_needed,
  total_cost = total_cost)

```
Questions

c)How many full-time English-speaking agents, full-time Spanish-speaking agents, and
part-time agents should Alex hire for each 2-hour shift to minimize operating costs
while attending to all calls? 

d)What is the minimum cost for the optimization model to assist Alex’s decision in
hiring all agents that she needs? 

```{r Part c,d, echo= FALSE, message=FALSE, comment = " " ,warning=FALSE}
# Define the hourly call forecast data
call_forecast <- c(40, 85, 70, 95, 80, 35, 10)

# Agent efficiency (calls per hour)
agent_efficiency <- 6

# Original full-time agents needed (sum of English and Spanish agents)
full_time_agents_needed <- ceiling(call_forecast / agent_efficiency / 2)

# Adjusting for the constraint: Only one English-speaking agent at 1 P.M. and 3 P.M.
full_time_english_agents_adjusted <- full_time_agents_needed
full_time_english_agents_adjusted[c(4, 5)] <- 1

# Calculating the number of Spanish-speaking agents needed
full_time_spanish_agents <- full_time_agents_needed - full_time_english_agents_adjusted

# Part-time agents needed
part_time_agents_needed <- c(ceiling(sum(call_forecast[5:6]) / agent_efficiency),
                             ceiling(sum(call_forecast[6:7]) / agent_efficiency))

# Wage rates
wage_before_5pm <- 30
wage_after_5pm <- 45

# Calculate the adjusted total cost
cost_full_time_adjusted <- sum(full_time_english_agents_adjusted[1:4] * wage_before_5pm * 2) +
                          sum(full_time_spanish_agents[1:4] * wage_before_5pm * 2) +
                          sum(full_time_english_agents_adjusted[5:7] * wage_after_5pm * 2) +
                          sum(full_time_spanish_agents[5:7] * wage_after_5pm * 2)
cost_part_time <- part_time_agents_needed[1] * wage_before_5pm * 4 +
                  part_time_agents_needed[2] * wage_after_5pm * 4
total_cost_adjusted <- cost_full_time_adjusted + cost_part_time

# Output the results
list(
  full_time_english_agents_adjusted = full_time_english_agents_adjusted,
  full_time_spanish_agents = full_time_spanish_agents,
  part_time_agents_needed = part_time_agents_needed,
  total_cost_adjusted = total_cost_adjusted
)
```
Questions

e)How many full-time and part-time agents should Alex hire for each 2-hour shift to
minimize operating costs while attending to all calls? 

```{r Part e, echo = FALSE, message=FALSE, comment = " " ,warning=FALSE}
# All agents are bilingual, so we don't need to separate them into languages
full_time_bilingual_agents_needed <- ceiling(call_forecast / agent_efficiency / 2)

list(full_time_bilingual_agents_needed = full_time_bilingual_agents_needed)

```
Questions

f)What is the minimum cost for the optimization model to assist Alex’s decision in hiring all agents that she needs? 

```{r Part f, echo = FALSE, message=FALSE, comment = " " ,warning=FALSE}
# minimum cost 
full_time_cost_before <-full_time_bilingual_agents_needed[0:4] * 30 
overtime_full_time_cost_before <- full_time_bilingual_agents_needed[5:6] * 45 

print(full_time_cost_before)
print(overtime_full_time_cost_before)
total <- sum(full_time_cost_before) + sum(overtime_full_time_cost_before)
print(total)
```
Questions

g) What is the maximum percentage increase in the hourly wage rate that Alex can offer to bilingual agents over monolingual agents without increasing the total operating costs?


```{r Part g, echo = FALSE, message=FALSE, comment = " " ,warning=FALSE}
# Define the hourly call forecast data
call_forecast <- c(40, 85, 70, 95, 80, 35, 10)

# Agent efficiency (calls per hour)
agent_efficiency <- 6

# Wage rates
wage_before_5pm <- 30
wage_after_5pm <- 45

# Calculate the cost for full-time bilingual agents
full_time_bilingual_agents_needed <- ceiling(call_forecast / agent_efficiency / 2)
cost_full_time_bilingual_before_5pm <- sum(full_time_bilingual_agents_needed[1:4] * wage_before_5pm * 2)
overtime_cost_full_time_bilingual_after_5pm <- sum(full_time_bilingual_agents_needed[5:7] * wage_after_5pm * 2)

# Calculate the cost for full-time monolingual agents
cost_full_time_monolingual_before_5pm <- sum(full_time_bilingual_agents_needed[1:4] * wage_before_5pm * 2)
overtime_cost_full_time_monolingual_after_5pm <- sum(full_time_bilingual_agents_needed[5:7] * wage_before_5pm * 2)

# Calculate the total cost for bilingual agents and monolingual agents
total_cost_bilingual <- cost_full_time_bilingual_before_5pm + overtime_cost_full_time_bilingual_after_5pm
total_cost_monolingual <- cost_full_time_monolingual_before_5pm + overtime_cost_full_time_monolingual_after_5pm

# Calculate the maximum percentage increase
max_percentage_increase <- ((total_cost_bilingual - total_cost_monolingual) / total_cost_monolingual) * 100
max_percentage_increase <- round(max_percentage_increase, 1)

# Output the maximum percentage increase
max_percentage_increase
```


Interpretation of Results
The optimization results indicate a strategic approach for Tesla to handle its customer service call volume efficiently. By aligning the number of agents (both full-time and part-time) with the varying call volumes throughout the day, Tesla can ensure high-quality customer support while minimizing idle time for agents. The introduction of bilingual agents aligns with Tesla’s commitment to diversity and inclusivity, potentially enhancing customer satisfaction and brand loyalty in a multilingual market like Silicon Valley.

Limitations
The primary limitation of this analysis is the reliance on several key assumptions, such as constant call handling efficiency and fixed shift durations. Real-world scenarios may present fluctuating call volumes, varying agent efficiency, and unexpected staffing needs. Additionally, the model does not account for potential changes in call patterns or the dynamic nature of customer service requirements over time.

Recommendations
Based on the findings, the following recommendations are proposed:

*Adaptable Staffing Strategies*: Develop a staffing strategy that is flexible and can adjust to changing call volumes and unexpected surges in demand.

*Multilingual Agent Recruitment*: Focus on recruiting agents who are fluent in multiple languages to better serve a diverse clientele, thereby improving communication and customer satisfaction.

*Ongoing Data Analysis*: Consistently evaluate call data to modify staffing plans according to evolving call trends.

*Agent Skills Enhancement*: Commit to the ongoing training of agents to enhance their efficiency and versatility, including their language skills and technical competencies.

*Feedback Implementation*: Set up systems to gather and utilize feedback from customers and agents to continually improve the customer service process.

*Utilizing Advanced Technology*: Employ advanced technology solutions for effective call routing, particularly for connecting callers with agents who have the appropriate language skills and subject matter expertise.

*Performance Monitoring and Adjustment*: Regularly assess the customer service department's performance using key indicators and make necessary strategic adjustments.

Conclusion

The analysis of Tesla's customer service strategy reveals a nuanced approach to staffing, emphasizing efficiency and responsiveness. Key findings suggest that a well-balanced mix of full-time, part-time, and bilingual agents, aligned with fluctuating call volumes, optimizes resource utilization while maintaining high service standards. The introduction of bilingual agents, catering to 20% non-English speaking customers, not only showcases Tesla's commitment to inclusivity but also significantly enhances customer engagement in a diverse market like Silicon Valley. This strategic approach underlines Tesla's dedication to exceptional service, a core aspect of its brand identity, and its broader mission of sustainable innovation.
